---
title: "ex3"
author: "Roy Gurwicz"
date: "11 11 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Doing DESeq2.
```{r}
library(ggplot2)
library(DESeq2)

counts_file <- system.file("extdata/rna-seq/SRP029880.raw_counts.tsv", package = "compGenomRData")
coldata_file <- system.file("extdata/rna-seq/SRP029880.colData.tsv", package = "compGenomRData")



counts <- read.table(counts_file,header = T, sep = '\t')
countData <- as.matrix(subset(counts,select=c(-width)))
colData <- read.table(coldata_file,header=T,sep='\t',stringsAsFactors = TRUE)

dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ group)

filteration <- rowSums(DESeq2::counts(dds)) >= 10
dds <- dds[filteration,]

dds <- DESeq(dds)

DEresults = results(dds,contrast = c("group",'CASE','CTRL'))
DEresultsDf <- as.data.frame(DEresults)
```

# 2. Use DESeq2::plotDispEsts to make a dispersion plot and find out the meaning of this plot. (Hint: Type ?DESeq2::plotDispEsts)

It is a plot of "Mean of (normalized) counts" vs. dispersion. This plot (in red) helps us decide the dispersion via the mean of how much a gene was counted.

```{r}
DESeq2::plotDispEsts(dds)
```


# 3. Explore lfcThreshold argument of the DESeq2::results function. What is its default value? What does it mean to change the default value to, for instance, 1?

```{r}
DEresults2Df = as.data.frame(DESeq2::results(dds,contrast = c("group",'CASE','CTRL'),lfcThreshold = 1))
```

It changes the threshold of the log2 fold change. 
The default value is 0.
Changing the default value to 1 the changes the Wald tests so that the p-value and the p-adj also change in acordance to the threshold.

# 4. What is independent filtering? 
# What happens if we don’t use it? 
# Google independent filtering statquest and watch the online video about independent filtering.
```{r}
DEresults3Df = as.data.frame(DESeq2::results(dds,contrast = c("group",'CASE','CTRL'),independentFiltering = FALSE))
#DEresults==DEresults3Df
#DEresultsDf[c("KRT33A"),]
#DEresults3Df(c("KRT33A"),)
```
It is a command which automatically filters genes with low read counts (for example, the gene KRT33A has NA result at its p-value, and therefore if we don't use the independentFiltering argument this gene is written in the data frame.

# 5. Re-do the differential expression analysis using the edgeR package. 
# Find out how much DESeq2 and edgeR agree on the list of differentially expressed genes.
They differ in 988 genes, but agree on 18,093 genes. DESeq2 has longer list of genes, which is encountered as differentially expressed, than edgeR.
```{r}
library(edgeR)
counts <- read.table(counts_file,header = T, sep = '\t')
countData <- as.matrix(subset(counts,select=c(-width)))
colData <- read.table(coldata_file,header=T,sep='\t',stringsAsFactors = TRUE)

ddsEdge <- DGEList(counts=countData, samples = colData, group = colData$group)

filterationEdge <- filterByExpr(ddsEdge)
ddsEdge <- ddsEdge[filterationEdge,]

ddsEdge <- calcNormFactors(ddsEdge)

resultsEdge <- model.matrix(~ colData$group)

# Estimates the dispersion.
ddsEdge <- estimateDisp(ddsEdge,resultsEdge) 
ddsEdgeDf <- as.data.frame(ddsEdge)

edgeCompare = as.data.frame(row.names(ddsEdge$counts))

DESeq2Compare = as.data.frame(row.names(DEresultsDf))

library(compare)
comparison <- compare(edgeCompare,DESeq2Compare,allowAll = T)$tM


differenceEdge <- lapply(1:ncol(edgeCompare),function(i)setdiff(edgeCompare[,i],comparison[,i]))
differenceEdgeDf <- data.frame(differenceEdge)

differenceDESeq <- lapply(1:ncol(DESeq2Compare),function(i)setdiff(DESeq2Compare[,i],comparison[,i]))
differenceDESeqDf <- data.frame(differenceDESeq)

summary(differenceDESeqDf)
summary(differenceEdgeDf)
```

# 6. Use the compcodeR package to run the differential expression analysis using at least three different tools 
# and compare and contrast the results following the compcodeR vignette.
```{r}
library(compcodeR)
setwd("C:/Users/roygu/Documents/Rcourse/3/")

# Creating a compData file.
colData$condition <- colData$group
info.parameters <- list(dataset = "mydata", uID = "123456")
cpd <- compData(countData,colData,info.parameters)

#cpdList <- convertcompDataToList(cpd)


generateSyntheticData("mydata.rds", 
                      n.vars = 19719, 
                      samples.per.cond = 5, 
                      n.diffexp = 19081, 
                      output.file = "mydata.rds")



# Run differential expression analysis.
runDiffExp(data.file = "mydata.rds",
           result.extent = "DESeq2", 
           Rmdfunction = "DESeq2.createRmd",
           output.directory = ".", 
           fit.type = "parametric", 
           test = "Wald", 
           beta.prior = TRUE, 
           independent.filtering = TRUE, 
           cooks.cutoff = TRUE, 
           impute.outliers = TRUE)

library(DSS)
runDiffExp(data.file = "mydata.rds", 
           result.extent = "DSS", 
           Rmdfunction = "DSS.createRmd",
           output.directory = ".", 
           norm.method = "quantile", 
           disp.trend = TRUE)

library(EBSeq)
runDiffExp(data.file = "mydata.rds", 
           result.extent = "EBSeq", 
           Rmdfunction = "EBSeq.createRmd",
           output.directory = ".", 
           norm.method = "median")



vignette("compcodeR")

library(rpanel)
runComparisonGUI(input.directories = ".", 
                 output.directory = ".", recursive = FALSE)



file.table <- data.frame(input.files =c("mydata_DESeq2.rds",
                                         "mydata_DSS.rds", 
                                        "mydata_EBSeq.rds"),
                         stringsAsFactors = FALSE)

runComparison(file.table = file.table, 
                parameters = NULL, 
                output.directory = ".")
```

